---
title: "Chapter 2 exercises"
output:
  github_document:
    toc: true
  html_document:
    toc: true
    df_print: paged
---

```{r}
library(fpp3)
```

# Ex 1

1. Use the help function to explore what the series gafa_stock, PBS, vic_elec and pelt represent.

    -  Use autoplot() to plot some of the series in these data sets.

    -  What is the time interval of each series?
      - gafa_stock is Business Daily
      - PBS is monthly
      - vic_elec is hourly (every other hour)
      - pelt is yearly
      
      
    
    
```{r}
gafa_stock
```
    
```{r}
gafa_stock |>
  autoplot(Close) +
  labs(title = "Closing price for various stocks",
       subtitle = "2014-2019",
       y = "")
```

```{r}
PBS
```

```{r}
PBS |> 
  select(Month, Cost) |>
  summarise(TotalCost = sum(Cost)) |>
  autoplot(TotalCost) +
    labs(title = "Cost of various pharmaceuticals",
       y = "")
```
```{r}
vic_elec
```
```{r}
vic_elec |>
  autoplot(Temperature) +
  labs(title = "Temperatures in Melborne (30 minute intervals)", y="")
```

```{r}
pelt
```

```{r}
pelt |>
  autoplot(Hare) +
  labs(title = "Annual hare pelts harvested",
       y = "Year")
```

```{r}
pelt |>
  autoplot(Lynx) +
  labs(title = "Annual lynx pelts harvested",
       y = "Year")
```

# Ex 2

2. Use filter() to find what days corresponded to the peak closing price for each of the four stocks in gafa_stock.
  - Google: 1268.33 on 7/26/18
  - Facebook: 217.50 on 7/25/18
  - Amazon: 2039.51 on 9/4/18
  - Apple: 232.07 on 10/3/18

```{r}
aapl_stock <- filter(gafa_stock, Symbol == "AAPL")
summary(aapl_stock[,"Close"])
```
```{r}
aapl_stock_max <- max(aapl_stock[,"Close"])
aapl_stock_max
```

```{r}
aapl_stock |>
  filter(Close == aapl_stock_max)
```

```{r}
amzn_stock <- filter(gafa_stock, Symbol == "AMZN")
max(amzn_stock[,"Close"]) -> amzn_stock_max
amzn_stock |>
  filter(Close == amzn_stock_max)
```
```{r}
fb_stock <- filter(gafa_stock, Symbol == "FB")
max(fb_stock[,"Close"]) -> fb_stock_max
fb_stock |>
  filter(Close == fb_stock_max)
```

```{r}
goog_stock <- filter(gafa_stock, Symbol == "GOOG")
max(goog_stock[,"Close"]) -> goog_stock_max
goog_stock |>
  filter(Close == goog_stock_max)
```

# Ex 3

Download the file `tute1.csv` from the [book website]("https://bit.ly/fpptute1"), open it in Excel (or some other spreadsheet application), and review its contents. You should find four columns of information. Columns B through D each contain a quarterly series, labelled Sales, AdBudget and GDP. Sales contains the quarterly sales for a small company over the period 1981-2005. AdBudget is the advertising budget and GDP is the gross domestic product. All series have been adjusted for inflation.

a. Read data inR

```{r}
tute1 <- readr::read_csv("tute1.csv")
tute1
```

b. Convert data to time series

```{r}
tute_ts <- tute1 |>
  mutate(Quarter = yearquarter(Quarter)) |>
  as_tsibble(index = Quarter)
tute_ts
```

c. Construct time series plots of each of the three series

```{r}
tute_ts |>
  pivot_longer(-Quarter) |>
  ggplot(aes(x = Quarter, y = value, colour = name)) +
  geom_line() +
  facet_grid(name ~ ., scales = "free_y")
```

# Exercise 4

The USgas package contains data on the demand for natural gas in the US.

Install the USgas package.
Create a tsibble from us_total with year as the index and state as the key.
Plot the annual natural gas consumption by state for the New England area (comprising the states of Maine, Vermont, New Hampshire, Massachusetts, Connecticut and Rhode Island).






















