---
title: "Chapter 5 The forecaster's toolbox"
output:
  github_document:
    toc: true
  html_document:
    toc: true
    df_print: paged
---

The feasts package includes functions for computing FEatures And Statistics from Time Series.

```{r}
library(fpp3)
library(feasts)
```

# 5.1 A tidy forecasting workflow

To illustrate the process, we will fit linear trend models to national GDP data stored in `global_economy`.

## Data preparation (tidy)

The first step in forecasting is to prepare data in the correct format. This process may involve loading in data, identifying missing values, filtering the time series, and other pre-processing tasks. The functionality provided by `tsibble` and other packages in the `tidyverse` substantially simplifies this step.

If we want to model GDP per capita over time we must first compute the variable

```{r}
gdppc <- global_economy |>
  mutate(GDP_per_capita = GDP / Population)
gdppc
```

## Visualize the data

Let's look at the data for one country

```{r}
gdppc |>
  filter(Country == "Sweden") |>
  autoplot(GDP_per_capita) +
  labs(y = "$US", title = "GDP per capita for Sweden")
```

## Specify a model

There are many different time series models that can be used for forecasting, and much of this book is dedicated to describing various models. Specifying an appropriate model for the data is essential for producing appropriate forecasts.

Models in fable are specified using model functions, which each use a formula (y ~ x) interface. The response variable(s) are specified on the left of the formula, and the structure of the model is written on the right.

For example, a linear trend model (to be discussed in Chapter 7) for GDP per capita can be specified with

```{r}
TSLM(GDP_per_capita ~ trend())
```

In this case the model function is `TSLM()` (time series linear model), the response variable is `GDP_per_capita` and it is being modelled using `trend()` (a “special” function specifying a linear trend when it is used within `TSLM()`). We will be taking a closer look at how each model can be specified in their respective sections.

The special functions used to define the model’s structure vary between models (as each model can support different structures). The “Specials” section of the documentation for each model function lists these special functions and how they can be used.

The left side of the formula also supports the transformations discussed in Section 3.1, which can be useful in simplifying the time series patterns or constraining the forecasts to be between specific values (see Section 13.3).

## Train the model (estimate)

Once an appropriate model is specified, we next train the model on some data. One or more model specifications can be estimated using the model() function.

```{r}
fit <- gdppc |>
  model(trend_model = TSLM(GDP_per_capita ~ trend()))
fit
```

This fits a linear trend model to the GDP per capita data for each combination of key variables in the `tsibble`. In this example, it will fit a model to each of the 263 countries in the dataset. The resulting object is a model table or a “**mable**”.

## Evaluate model performance

Once a model has been fitted, it is important to check how well it has performed on the data. There are several diagnostic tools available to check model behaviour, and also accuracy measures that allow one model to be compared against another. Sections 5.8 and 5.9 go into further details.

## Produce forecasts

With an appropriate model specified, estimated and checked, it is time to produce the forecasts using forecast(). The easiest way to use this function is by specifying the number of future observations to forecast. For example, forecasts for the next 10 observations can be generated using h = 10. We can also use natural language; e.g., h = "2 years" can be used to predict two years into the future.

In other situations, it may be more convenient to provide a dataset of future time periods to forecast. This is commonly required when your model uses additional information from the data, such as exogenous regressors. Additional data required by the model can be included in the dataset of observations to forecast.

```{r}
fit |> forecast(h = "3 years")
```

This is a forecast table, or “**fable**”. Each row corresponds to one forecast period for each country. The `GDP_per_capita` column contains the **forecast distribution**, while the `.mean` column contains the **point forecast**. The point forecast is the mean (or average) of the forecast distribution.

The forecasts can be plotted along with the historical data using autoplot() as follows.

```{r}
fit |>
  forecast(h = "3 years") |>
  filter(Country == "Sweden") |>
  autoplot(gdppc) +
  labs(y = "$US", title = "GDP per capita for Sweden")
```

# 5.2 Some simple forecasting methods

Some forecasting methods are extremely simple and surprisingly effective. We will use four simple forecasting methods as benchmarks throughout this book. 

To illustrate them, we will use quarterly Australian clay brick production between 1970 and 2004.

The `filter_index()` function is a convenient shorthand for extracting a section of a time series.

```{r}
bricks <- aus_production |>
  filter_index("1970 Q1" ~ "2004 Q4") |>
  select(Bricks)
bricks
```

## Mean method

Here, the forecasts of all future values are equal to the average (or “mean”) of the historical data. If we let the historical data be denoted by $y_1,\dots,y_T$, then we can write the forecasts as


$$\hat{y}_{T+h|T}=\bar{y}=(y_1+\dots+y_T)/T.$$
The notation $\hat{y}_{T+h|T}$ is a short-hand for the estimate of $y_{T+h}$ based on the data $y_1,\dots,y_T$

```{r}
bricks |> model(MEAN(Bricks))
```

## Naive method

All forecasts set to the value of the last observation.

$$\hat{y}_{T+h|T} = y_{T}$$
 
 This method works well for many economic/financial forecasts.
 
```{r}
bricks |> model(NAIVE(Bricks))
```
 
## Seasonal naive method

A similar method is useful for highly seasonal data. In this case, we set each forecast to be equal to the last observed value from the same season (e.g., the same month of the previous year). Formally, the forecast for time $T+h$ is written as

$$\hat{y}_{T+h|T} = y_{T+h-m(k+1)}$$

where $m=$ the seasonal period, and $k$ is the integer part of $(h−1)/m$ (i.e., the number of complete years in the forecast period prior to time $T+h$). This looks more complicated than it really is. For example, with monthly data, the forecast for all future February values is equal to the last observed February value. With quarterly data, the forecast of all future Q2 values is equal to the last observed Q2 value (where Q2 means the second quarter). Similar rules apply for other months and quarters, and for other seasonal periods.

```{r}
bricks |> model(SNAIVE(Bricks ~ lag("year")))
```

The `lag()` function is optional here as bricks is quarterly data and so a seasonal naïve method will need a one-year lag. However, for some time series there is more than one seasonal period, and then the required lag must be specified.



## Drift method

A variation on the naïve method is to allow the forecasts to increase or decrease over time, where the amount of change over time (called the **drift**) is set to be the average change seen in the historical data. Thus the forecast for time $T+h$ is given by

$$\hat{y}_{T+h|T} = y_{T} + \frac{h}{T-1}\sum_{t=2}^T (y_{t}-y_{t-1}) = y_{T} + h \left( \frac{y_{T} -y_{1}}{T-1}\right)$$

This is equivalent to drawing a line between the first and last observations, and extrapolating it into the future.

```{r}
bricks |> model(RW(Bricks ~ drift()))
```

## Example: Australian quarterly beer production

The first three methods applied to Australian quarterly beer production from 1992 to 2006, with the forecasts compared against actual values in the next 3.5 years.

```{r}
# Set training data from 1992 to 2006
train <- aus_production |>
  filter_index("1992 Q1" ~ "2006 Q4")
# Fit the models
beer_fit <- train |>
  model(
    Mean = MEAN(Beer),
    `Naïve` = NAIVE(Beer),
    `Seasonal naïve` = SNAIVE(Beer)
  )
# Generate forecasts for 14 quarters
beer_fc <- beer_fit |> forecast(h = 14)
# Plot forecasts against actual values
beer_fc |>
  autoplot(train, level = NULL) +
  autolayer(
    filter_index(aus_production, "2007 Q1" ~ .),
    colour = "black"
  ) +
  labs(
    y = "Megalitres",
    title = "Forecasts for quarterly beer production"
  ) +
  guides(colour = guide_legend(title = "Forecast"))
```
 
In this case, only the seasonal naïve forecasts are close to the observed values from 2007 onwards

## Example: Google's daily closing stock price

The non-seasonal methods are applied to Google’s daily closing stock price in 2015, and used to forecast one month ahead. Because stock prices are not observed every day, we first set up a **new time index based on the trading days rather than calendar days**.

### Re-index based on trading days

```{r}
google_stock <- gafa_stock |>
  filter(Symbol == "GOOG", year(Date) >= 2015) |>
  mutate(day = row_number()) |>
  update_tsibble(index = day, regular = TRUE)
google_stock
```

### Filter the year of interest

```{r}
google_2015 <- google_stock |> filter(year(Date) == 2015)
google_2015
```

### Fit the models

```{r}
google_fit <- google_2015 |>
  model(
    Mean = MEAN(Close),
    `Naïve` = NAIVE(Close),
    Drift = NAIVE(Close ~ drift())
  )
google_fit
```

### Produce forecasts for the trading days in January 2016

```{r}
google_jan_2016 <- google_stock |>
  filter(yearmonth(Date) == yearmonth("2016 Jan"))
google_fc <- google_fit |>
  forecast(new_data = google_jan_2016)
```

### Plot the forecasts

```{r}
google_fc |>
  autoplot(google_2015, level = NULL) +
  autolayer(google_jan_2016, Close, colour = "black") +
  labs(y = "$US",
       title = "Google daily closing stock prices",
       subtitle = "(Jan 2015 - Jan 2016") +
  guides(colour = guide_legend(title = "Forecast"))
```

Sometimes one of these simple methods will be the best forecasting method available; but in many cases, these methods will serve as **benchmarks** rather than the method of choice. That is, any forecasting methods we develop will be compared to these simple methods to ensure that the new method is better than these simple alternatives. If not, the new method is not worth considering.

# 5.3 Fitted values and residuals

# 5.4 Residual diagnostics

# 5.5 Distributional forecasts and predictions

# 5.6 Forecasting using transformations

# 5.7 Forecasting with decomosition

# 5.8 Evaluating point forecast accuracy

# 5.9 Evaluating distributional forecast accuracy

# 5.10 Time series cross-validation






