---
title: "Chapter 4 Time series features"
output:
  github_document:
    toc: true
  html_document:
    toc: true
    df_print: paged
---

The feasts package includes functions for computing FEatures And Statistics from Time Series.

```{r}
library(fpp3)
library(feasts)
```


# 4.1 Some simple statistics

**Features** are numerical summmaries computed from a time series.`features()` is used in R for this purpose.

```{r}
tourism |>
  features(Trips, list(mean = mean, sd = sd)) |>
  arrange(mean)
```

This shows that Kangaroo Island/Other had the least average number of visits.

Using `quantile` in the `features` argument of `features()` provides the basic statistical information.

```{r}
tourism |> features(Trips, quantile)
```

# 4.2 ACF features

Autocorrelations were discussed in Section 2.8. All the autocorrelations of a series can be considered features of that series. We can also summarise the autocorrelations to produce new features; for example, the sum of the first ten squared autocorrelation coefficients is a useful summary of how much autocorrelation there is in a series, regardless of lag.

We can also compute autocorrelations of the changes in the series between periods. That is, we “difference” the data and create a new time series consisting of the differences between consecutive observations. Then we can compute the autocorrelations of this new differenced series. Occasionally reiteration of differencing may provide more useful information.

For example we may compute seasonal differences for monthly data, eg. Feb to Feb. This enables us to look at how the series is changing between years, rather than between months. Again, the autocorrelations of the seasonally differenced series may provide useful information.

The `feat_acf()` function computes a selection of the autocorrelations discussed here. It will return six or seven features:

- the first autocorrelation coefficient from the original data;
- the sum of squares of the first ten autocorrelation coefficients from the original data;
- the first autocorrelation coefficient from the differenced data;
- the sum of squares of the first ten autocorrelation coefficients from the differenced data;
- the first autocorrelation coefficient from the twice differenced data;
- the sum of squares of the first ten autocorrelation coefficients from the twice differenced data;
- For seasonal data, the autocorrelation coefficient at the first seasonal lag is also returned.

```{r}
tourism |> features(Trips, feat_acf)
```

# 4.3 STL features

The STL decomposition provides several more features. These can help identify the series with the most trend or seasonality among a large collection of time series.

For strongly trended data, the seasonally adjusted data should have much more variation than the remainder component. Therefore $Var(R_t)/Var(T_t+R_t)$ should be relatively small. But for data with little or no trend, the two variances should be approximately the same. So we define the **strength of trend** as:

$$
F_T = \max\left(0, 1 - \frac{\text{Var}(R_t)}{\text{Var}(T_t+R_t)}\right)
$$
$F_T$ has a range of [0,1].



Similarly the **strength of seasonality** is

$$
F_S = \max\left(0, 1 - \frac{\text{Var}(R_t)}{\text{Var}(S_{t}+R_t)}\right)
$$

```{r}
tourism |> features(Trips, feat_stl)
```

```{r}
tourism |> 
  features(Trips, feat_stl) |>
  ggplot(aes(x = trend_strength,
             y = seasonal_strength_year,
             col = Purpose)) +
  geom_point() +
  facet_wrap(vars(State))
```

This shows strong seasonal strength for "Holiday" and strong trend strength for Victoria and Western Australia.

The most seasonal series can be easily identified and plotted.

```{r}
tourism |>
  features(Trips, feat_stl) |>
  filter(
    seasonal_strength_year == max(seasonal_strength_year)
  ) |>
  left_join(tourism, 
            by = c("State", "Region", "Purpose"), 
            multiple = "all") |>
  ggplot(aes(x = Quarter, y = Trips)) +
  geom_line() +
  facet_grid(vars(State, Region, Purpose))
```

This is apparently the most popular ski region of Australia.

The `feat_stl()` function returns several more features other than those discussed above.

- `seasonal_peak_year` indicates the timing of the peaks — which month or quarter contains the largest seasonal component. This tells us something about the nature of the seasonality. In the Australian tourism data, if Quarter 3 is the peak seasonal period, then people are travelling to the region in winter, whereas a peak in Quarter 1 suggests that the region is more popular in summer.
- `seasonal_trough_year` indicates the timing of the troughs — which month or quarter contains the smallest seasonal component.
- `spikiness` measures the prevalence of spikes in the remainder component $R_t$ of the STL decomposition. It is the variance of the leave-one-out variances of $R_t$.
- `linearity` measures the linearity of the trend component of the STL decomposition. It is based on the coefficient of a linear regression applied to the trend component.
- `curvature` measures the curvature of the trend component of the STL decomposition. It is based on the coefficient from an orthogonal quadratic regression applied to the trend component.
- `stl_e_acf1` is the first autocorrelation coefficient of the remainder series.
- `stl_e_acf10` is the sum of squares of the first ten autocorrelation coefficients of the remainder series.

# 4.4 Other features

# 4.5 Exploring Australian tourism data