---
title: "Chapter 3 Time series decomposition"
output:
  github_document:
    toc: true
  html_document:
    toc: true
    df_print: paged
---

```{r}
library(fpp3)
```

We generally consider three types of patterns in time series:

- trend
- seasonality
- cycles

trend and cycle are usually combined into a single trend-cycle component often just called trend.

A time series can then be decomposed into:

- trend component
- seasonality component of which there may be multiple
- remainder component

# 3.1 Transformations and Adjustments

When decomposing a time series, it is sometimes helpful to first transform or adjust the series in order to make the decomposition (and later analysis) as simple as possible. So we will begin by discussing transformations and adjustments.

## Calendar adjustments

Use average/month rather than raw totals

## Population adjustments

Convert to per capita data where appropriate.

```{r}
global_economy |>
  filter(Country == "Australia") |>
  autoplot(GDP/Population) +
  labs(title = "GDP per capita", y = "$US")
```

## Inflation adjustments

The Consumer Price Index (CPI) can be used to adjust current values for inflation based on a prior time. The price index $z_t$ and the original hose price in year $t$ as $y_t$ can be used to calculated the adjusted value by $x_t=t_t/z_t*z_{2000}$ which would give the adjusted price based on year 2000 dollar values.

As an example, we can look at the annual newspaper and book sales in `aus_retail` and adjust the data for inflation using CPI from `global_economy`


```{r}
print_retail <- aus_retail |>
  filter(Industry == "Newspaper and book retailing") |>
  group_by(Industry) |>
  index_by(Year = year(Month)) |>
  summarise(Turnover = sum(Turnover))
aus_economy <- global_economy |>
  filter(Code == "AUS")
```

```{r}
print_retail |>
  left_join(aus_economy, by = "Year") |>
  mutate(Adjusted_turnover = Turnover / CPI * 100) |>
  pivot_longer(c(Turnover, Adjusted_turnover),
               values_to = "Turnover") |>
  mutate(name = factor(name,
         levels=c("Turnover","Adjusted_turnover"))) |>
  ggplot(aes(x = Year, y = Turnover)) +
  geom_line() +
  facet_grid(name ~ ., scales = "free_y") +
  labs(title = "Turnover: Australian print media industry",
       y = "$AU")
```

By adjusting for inflation we see that the industry is in steeper decline than the raw data suggests.

## Mathematical transformations

**Logarithmic transformations** $w_t=\log({y_t})$, are useful when variation increases and decreases with the level of the series.

**Power transformations** $w_t=y_t^p$ are also used but are more difficult to interpret.

**Box-Cox transformations**

$$
\begin{equation}
  w_t  =
    \begin{cases}
      \log(y_t) & \text{if $\lambda=0$};  \\
      (\text{sign}(y_t)|y_t|^\lambda-1)/\lambda & \text{otherwise}.
    \end{cases}
\end{equation}
$$

In the modified Box-Cox used here negative values of $y_t$ are allowed as long as $\lambda>0$ 

The logarithm used is $ln$.

$\lambda$ should be chosen to make the seasonal variation relatively constant accross the series.

`gerrero` can be used to determine $\lambda$.

```{r}
lambda <- aus_production |>
  features(Gas, features = guerrero) |>
  pull(lambda_guerrero)
lambda
```
```{r}
# install.packages("latex2exp")
aus_production |>
  autoplot(box_cox(Gas, lambda)) +
  labs(y = "",
       title = latex2exp::TeX(paste0(
         "Transformed gas production with $\\lambda$ = ",
         round(lambda,2)
       )))
```




